### 2023-04-28 chapter2/chapter2.tex
Finished paragraphs on contrast analyses
Added contrast analyses paragraphs to discussion

% III. Disorder score analyses
%    A. The analysis of feature and score contrasts is effectively a natural ``mutational scanning'' experiment. Though the mutations generated by evolution are not a random or exhaustive sample of sequence space, they are perturbations around a common ancestral sequence which can reveal the relationship between a region's biophysical properties and its propensity for disorder. Accordingly, we find disorder scores have the strongest correlations with features that measure a region's overall hydrophobicity, \textit{e.g.} fraction_disorder and hydropathy. As these features are known correlates of intrinsic disorder, this relationship is expected. However, the strength of these correlations indicate that a region's relative proportion of hydrophilic and hydrophobic residues is, to a first approximation, the largest determinant of its structural state.
%        3. Excluding these hydrophobicity-related features, the second strongest association is a negatively correlation with wf_complexity, indicating the predictor strongly associates low complexity with intrinsic disorder. However, sequence complexity is a statistical rather than a biophysical criterion, and while many disordered regions have low levels of complexity, some low-complexity regions, like collagen, are structured. This suggests that while disorder predictors are in general accurate classifiers of a residue's structural state, they can conflate correlation and causation, and therefore in some cases their predictions may require careful interpretation.
%
%        5. Interestingly, the disorder scores have significant, if weak, correlations with most features used in this analysis, including the motifs. Disorder scores are weakly correlated with many motifs. Furthermore, unclear if predictor is responding directly to these motifs as a signal of disorder, or they appear in regions which evolving towards disorder. Interestingly docking (DOC) and ligand binding (LIG) sites tend to have negative correlations whereas the modification (MOD) sites have positive correlations. Docking and ligand binding sites are generally mediated by small hydrophobic patches, so the predictor could be responding to increased hydrophobicity caused by additional of hydrophobic residue ``completing'' the motif rather than the motif itself. Likewise, phosphorylation sites are highly enriched in IDRs and many are targeted to disorder-promoting residues like serine or threonine. Thus, it's unclear if the predictor is responding directly to these features or features which are correlated with them. More in depth analyses of the representations of these disorder predictors would shed led on these questions.
%        1. In, the disorder scores have a significant relationship with most features used this analysis.
%        4. Furthermore, resolution of analyses limited because features calculated at level of regions, whose lengths can exceed 1,000 residues, rather than a fixed window around each residue.

%    B. The GO annotation enrichment analysis indicate that proteins involved in DNA repair and the formation of extracellular structure contain disproportionate number of regions whose disorder scores are rapidly evolving. Because the significance tests were not corrected for multiple testing or controlled for false discovery rate, we caution against over-interpreting this result and instead consider as a hypothesis for further investigation. In general, however, these regions with rapidly evolving disorder scores may correspond to molecular recognition features (MoRFs), which are modules in IDRs that undergo a disorder-to-order transition on binding their targets. Because MoRFs already exist on the boundary between disorder and structure, small changes in the biophysical properties of these regions may have disproportionate effects on their structural ensembles. Thus, the most variable disorder scores may reflect instances where a mutation triggered a ``phase transition'' between largely structured or disordered native state.

@@

\section{Discussion}
Here's where my discussion will go!

>>>>

@@

\section{Discussion}
The analysis of feature and score contrasts is effectively a natural ``mutational scanning'' experiment. Though the mutations generated by evolution are not a random or exhaustive sample of sequence space, they are perturbations of a common ancestor which can reveal the relationship between a region's biophysical properties and its propensity for disorder. Accordingly, we find disorder scores have the strongest correlations with features that measure a region's overall polarity and hydrophobicity, \textit{e.g.} fraction\_disorder and hydropathy. As the formation of a hydrophobic core is a major driving force in protein folding, this relationship is expected. However, the strength of these correlations indicate that a region's relative proportion of hydrophilic and hydrophobic residues is, to a first approximation, the largest determinant of intrinsic disorder.

Excluding these hydrophobicity-related features, the next strongest association is a negative correlation with wf\_complexity, demonstrating the predictor strongly associates low complexity with intrinsic disorder. However, sequence complexity is a statistical rather than a biophysical criterion, and while many disordered regions have low levels of complexity, some low-complexity regions, like collagen, are structured. This suggests that while disorder predictors are in general accurate classifiers of a residue's structural state, they can conflate the correlates of intrinsic disorder with their causes. Therefore, in some cases their predictions may require careful interpretation.

The remaining significant correlations are generally weak and likely reflect a partial redundancy with the more strongly correlated features discussed previously. Interestingly, though, disorder scores are weakly correlated with many motifs, with the signs largely reflecting their class. For example, the correlations with docking (DOC) and ligand binding (LIG) sites are largely negative whereas those with modification (MOD) sites are generally positive. However, this analysis does not indicate whether the predictor responds to these motifs directly or to features that are correlated with them. For example, docking and ligand binding sites are generally mediated by small hydrophobic patches, so the correlations could reflect an increase in hydrophobicity caused by an additional hydrophobic residue ``completing'' the motif rather than the motif itself. Likewise, IDRs are highly enriched in phosphorylation sites, many of which are targeted to disorder-promoting residues like serine or threonine. As the disorder scores and features are calculated at the level of regions, whose lengths can exceed 1,000 residues, this analysis is limited in its ability to distinguish these possibilities. However, a more targeted \textit{in silico} mutational analysis would yield further insights.

The GO annotation enrichment analysis indicates that proteins involved in DNA repair and extracellular structures contain a disproportionate number of regions whose disorder scores are rapidly evolving. Because the significance tests were not corrected for multiple testing or controlled for their false discovery rate, we caution against over-interpreting this result and instead consider it as a hypothesis for further investigation. In general, however, these regions with rapidly evolving disorder scores may correspond to molecular recognition features (MoRFs), which are modules in IDRs that undergo a disorder-to-order transition on binding their targets. Because MoRFs already exist on the boundary between disorder and structure, small changes in the biophysical properties of these regions may have disproportionate effects on their structural ensembles. Thus, the most variable disorder scores may reflect instances where a mutation triggered a ``phase transition'' between largely structured or disordered native states.
